Transparency to hybrid open access invoicing through publisher-provided metadata: A longitudinal article-level study of Elsevier

### 2021-01-29 analysis/manuscript.Rmd
version pushed to overleaf for final polishing

The problem with this lack of transparency becomes more apparent in the case of hybrid OA. In this OA business model, individual articles can be made openly available for a fee, also known as article-processing charge (APC), while the journal as a whole remains subscription access. Hence, hybrid OA was originally introduced as a low-risk transitional model that allows journals to gradually convert to full OA and reduce subscription costs as the OA uptake increases [@Prosser_2003]. Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years. This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge. However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015]. While publishers assure they combine subscription fees and APCs to adjust pricing [@Mittermaier_2015], without transparency around the uptake of hybrid OA and revenue streams, such claims are impossible to evaluate.

With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased. If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017]. Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018]. However, as formal bibliometrics and financial studies have noted, the lack of publicly available and standardized data about hybrid OA publishing is considerable [@Laakso_2016; @Lawson_2015; @Pinfield_2016]

This current study investigates the extent to which publisher-provided data can be utilized to assess hybrid OA. The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing. Our approach is based on openly available data sources and brings together scholarly data from Crossref and machine-readable APC invoicing data from Elsevier OA full-texts, which enables us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program. Further, we investigate whether hybrid publishing fees were waived or charged to the authors or academic institutions or research funders with Elsevier OA agreements. In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject. As such, our findings have implications for research and OA policy and highlight the potential of publisher-provided data for large-scale, instantaneous analysis of hybrid OA uptake and invoicing.

@@

Other studies assessed the share of (hybrid) OA for the entire corpus of scholarly research articles. For instance, using research articles with an assigned DOI and indexed in the Web of Science Core Collection (n=2,610,305) as a benchmark, @Mart_n_Mart_n_2018 assessed how many research articles from 2009 and 2014 are freely available through Google Scholar. After identifying OA articles through licensing information from Crossref, the authors estimated the share of hybrid OA at around 0.5% in 2009 and 1.5% in 2014. In a similar vein, @Piwowar_2018 drew on Crossref to identify content licenses. The authors followed a data-driven approach based on Unpaywall data, which is openly available and contained 66 million Crossref DOIs at the time, to determine how much of the scholarly literature is OA. Based on a random sample of 100,000 Crossref DOIs, the authors found 3.6% of hybrid OA, while pointing towards a growing trend in recent years that recorded a hybrid OA share of 9.4% of all articles published in 2015. While Piwowar et al.'s (2018) and Martín-Martín et al.'s (2018) methods are well-suited for large-scale and repeated analyses of the prevalence of hybrid OA, their main limitation is that they lack information about publishing costs.

Moreover, Unpaywall's hybrid OA detection method has been refined several times since its launch in 2017, reflecting growing albeit incomprehensive license information in Crossref metadata. To overcome these shortcomings, the developers extended previous web-scraping approaches to better distinguish between immediate and delayed OA based on licenses at the date of observation @Piwowar_2019. Relatedly, in October 2020, Elsevier's license for delayed OA content was added to Unpaywall's list of license patterns [@Unpaywall_oa_license]. Finally, at the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types]. Because of these ongoing changes due to non-standardized data, uptake studies reporting hybrid OA uptake based on Unpaywall data cannot easily be compared with each other.

@@

As this brief review has shown, the most comprehensive studies approach hybrid OA funding from an institutional perspective that focuses on research institutions or funders and draws on internal records or crowd-sourced data like from the Open APC initiative. The advantages of Open APC data are that they reflect the actual APCs paid rather than list prices; they are not limited to certain publishers or journals; they are provided in a standardized format regardless of the publisher. On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity. Moreover, since Open APC data does not include waivers or free APCs, it cannot present a complete picture of all publishing activities and expenses. This, however, would pose less of an issue when analyzing just hybrid OA.

## Data-driven workflows for hybrid OA

In September 2018, coalition S presented a set of principles known as Plan S and announced. that from 2021 onwards, signatories would no longer fund publications in hybrid OA journals. This principle includes so-called "mirror" and sister journals—hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019]. While the scale and means of these transformative agreements may differ, many address the lack of transparency surrounding hybrid OA by making publisher-provided metadata a defining element of publishing workflows between institutions and publishers [@Geschuhn_2017; @Marques_2019]

Elsevier, one of the world's largest scholarly publishers, provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes. Surprisingly, besides the full-text article, the XML files also contain invoicing information and the article's OA status. Table \ref{tab:els_xml} summarizes the key metadata elements, which comprise the type of entity to which the APC was invoiced (`openaccessSponsorType`) and the name of that entity if available (`openaccessSponsorName`). Moreover, the article-level metadata indicates if OA was provided through Elsevier's Open Archive program (`openArchiveArticle`), which enables free access to participating subscription journals after an embargo period (i.e., delayed OA; @Els_Archive). As we have been unable to identify similar efforts by other large publishers (e.g., Springer Nature), this increased transparency through openly available bibliometric and financial data at the article level seems unique to Elsevier.

\begin{table}[H]
\caption{\label{tab:els_xml}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
\centering
\begin{tabular}[t]{lp{10cm}}
\toprule
XML Node & Description\\
\midrule
\texttt{openaccessArticle} & Indicates if the article is open
access. \\
\texttt{openaccessSponsorType} & Invoice recipient type.\\
\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
invoice.\\
\texttt{openArchiveArticle} & Indicates if open access was provided through
Elsevier's Open Archive programme.\tabularnewline \\
\bottomrule
\end{tabular}
\end{table}

Table 1: Overview of metadata provided in XML-based Elsevier full-texts representing information about open access status and invoicing. A sample XML is available here: https://api.elsevier.com/content/article/PII:S0169409X18301479?httpAccept=text/xml

@@

>>>>

The problem with this lack of transparency becomes more apparent in the case of hybrid OA. In this OA business model, individual articles can be made openly available for a fee, also known as article-processing charge (APC), while the journal as a whole remains subscription access. Hence, hybrid OA was originally introduced as a low-risk transitional model that allows journals to gradually convert to full OA and reduce subscription costs as the OA uptake increases [@Prosser_2003]. Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years. This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge [@Laakso_2016]. However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015]. While publishers assure they combine subscription fees and APCs to adjust pricing [@Mittermaier_2015], without transparency around the uptake of hybrid OA and revenue streams, such claims are impossible to evaluate.

With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased. If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017; @Schimmer_2015]. Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018]. However, as formal bibliometrics and financial studies have noted, the lack of publicly available and standardized data about hybrid OA publishing is considerable [@Laakso_2016; @Lawson_2015; @Pinfield_2016]

This current study investigates the extent to which publisher-provided data can be utilized to research hybrid OA. The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing. Our approach is based on openly available data sources and brings together scholarly metadata from Crossref, a DOI registration agency for scholarly works, and machine-readable APC invoicing data from Elsevier OA full-texts. The so-compiled data enable us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program. Further, we investigate whether hybrid publishing fees were waived or charged. In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject. As such, our findings have implications for research and OA policy and highlight the potential of publisher-provided data for large-scale, instantaneous analysis of hybrid OA uptake and invoicing.

@@

Other studies assessed the share of (hybrid) OA for the entire corpus of scholarly research articles. For instance, using research articles with an assigned DOI and indexed in the Web of Science Core Collection (n=2,610,305) as a benchmark, @Mart_n_Mart_n_2018 assessed how many research articles from 2009 and 2014 are freely available through Google Scholar. After identifying OA articles through licensing information from Crossref, the authors estimated the share of hybrid OA at around 0.5% in 2009 and 1.5% in 2014. In a similar vein, @Piwowar_2018 drew on Crossref to identify content licenses. The authors followed a data-driven approach based on Unpaywall data, which is openly available and contained 66 million Crossref DOIs at the time, to determine how much of the scholarly literature is OA. Based on a random sample of 100,000 Crossref DOIs, the authors found 3.6% of hybrid OA, while pointing towards a growing trend in recent years that recorded a hybrid OA share of 9.4% of all articles published in 2015.

Since its launch, Unpaywall has become widely used as an OA evidence source in bibliometric research and rankings[@Robinson_Garcia_2020; @Huang_2020]. However, results cannot easily be compared with each other over time because of Unpaywall's ongoing changes due to non-standardized data to discover hybrid OA. These changes mainly reflect the growing albeit incomprehensive license information in Crossref metadata. To overcome these shortcomings, the developers extended previous web-scraping approaches to better distinguish between immediate and delayed OA based on licenses at the date of observation @Piwowar_2019. Relatedly, in October 2020, Elsevier's license for delayed OA content was added to Unpaywall's list of license patterns [@Unpaywall_oa_license]. At the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types]. With its servie unsub, Unpaywall recently launched a paid service, where libraries can compare hybrid OA uptake rates with internal financial and usage data to assess the value of journal subscriptions [@Chawla_2020].

@@

As this brief review has shown, the most comprehensive studies approach hybrid OA funding from an institutional perspective that focuses on research institutions or funders and draws on internal records or crowd-sourced data like from the Open APC initiative. The advantages of Open APC data are that they reflect the actual APCs paid rather than list prices; they are not limited to certain publishers or journals; they are provided in a standardized format regardless of the publisher. On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity [@Pieper_2018]. Moreover, since Open APC data does not include waivers or free APCs, it cannot present a complete picture of all publishing activities and expenses. This, however, would pose less of an issue when analyzing just hybrid OA.

## The role of publisher-provided data

Uptake and financial studies have OA policy implications for hybrid OA. Most prominently, the cOAlition S, an international network of research funders, released its widely discussed Plan S. In September 2018, the cOAlition S announced that from 2021 onwards, signatories would no longer fund articles in hybrid OA journals. This principle includes so-called "mirror" and sister journals--hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019]. The Plan S refers to the Efficiency and Standards for Article Charges (ESAC) guidelines [@Geschuhn_2017] as critical component of transformative agreements, emphasizing the need for publisher-provided metadata to address the lack of transparency surrounding hybrid OA. However, studies confirm that publisher with such agreements lack essential data about hybrid OA uptake and invoicing, or are not able or willing to provide such information publicly and instantaneously [@Monaghan_2020; @Marques_2019], which, in turn, not just limit the monitoring of hybrid OA, but also research studies on its size and financial implications.

Here, we focus on Elsevier. The publisher provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes. Surprisingly, besides the full-text article, the XML files also contain invoicing information and the article's OA status. Table \ref{tab:els_xml} summarizes the key metadata elements, which comprise the type of entity to which the APC was invoiced (`openaccessSponsorType`) and the name of that entity if available (`openaccessSponsorName`). Moreover, the article-level metadata indicates if OA was provided through Elsevier's Open Archive program (`openArchiveArticle`), which enables free access to participating subscription journals after an embargo period (i.e., delayed OA; @Els_Archive). As we have been unable to identify similar efforts by other large publishers (e.g., Springer Nature), this increased transparency through openly available bibliometric and financial data at the article level seems unique to Elsevier.

\begin{table}[H]
\caption{\label{tab:els_xml} Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019}
\centering
\begin{tabular}[t]{lp{10cm}}
\toprule
XML Node & Description\\
\midrule
\texttt{openaccessArticle} & Indicates if the article is open
access. \\
\texttt{openaccessSponsorType} & Invoice recipient type.\\
\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
invoice.\\
\texttt{openArchiveArticle} & Indicates if open access was provided through
Elsevier's Open Archive programme.\tabularnewline \\
\bottomrule
\end{tabular}
\end{table}

@@

# Discussion

Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%. As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas. In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}). This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles. This order mostly reflects the disciplines' overall publication output. According to the Open Science Monitor by the European Commission, the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.

Such disciplinary differences in hybrid OA prevalence become more meaningful when considering the relative share of OA articles to closed-access articles. In line with previous research, we found that Elsevier journals from the life sciences and social sciences [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded greater than typical hybrid OA uptake, whereas physical sciences journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016]. In a systematic review of OA publishing patterns, @Severin_2020 linked such disciplinary differences to socio-cultural and technological factors that shape publishing cultures and practices. Among other aspects, the authors pointed to the availability of prestigious journals, the possibility to provide OA through alternative routes, and ease of access to funding sources to cover publishing fees. As such, looking at hybrid OA uptake in isolation can only provide limited insight, so that placing it within the broader context of disciplinary publishing and OA practices is necessary. For instance, since many branches of the physical sciences have established self-archiving practices and the technical infrastructure to support these (e.g., the arXiv; @Severin_2020, @Bj_rk_2010, @Laakso_2016), researchers can provide OA through repositories (i.e., green OA), and indeed, according to the Open Science Monitor the physical sciences provide OA primarily through the green route (31.2% total OA share, 24% OA if only considering green OA). Hence, the low share of hybrid OA in our sample might be because physical scientists perceive much less of a need for hybrid OA. Further, @Severin_2020 note project-based funding structures enable OA publishing in fully OA and hybrid journals as they allow for easy integration of publishing costs. Since these funding structures are typical for the life sciences (ibid.), we would expect higher rates of OA. Data from the Open Science Monitor seems to support this notion as the life sciences record particularly high shares of full OA articles and comparatively high  shares of hybrid OA in some fields, such as biology and agricultural science. While we expected the observed hybrid OA shares for the life sciences and physical sciences, we were surprised by the high share of hybrid OA belonging to the AJSC subject area social sciences, which also includes the humanities, since articles "social scientists have reported to face significant difficulties in access to grant funding for [publication charges], as most research in the social sciences is not done by means of project-specific funding" @Severin_2020 [p. 15]. However, previous research has found comparatively low awareness and uptake of green OA among researchers in the  field [@Bj_rk_2010; @Creaser_2010], which might encourage hybrid OA uptake. Another possible explanation,is that Elsevier has a number of high prestigious social science and humanities journals [@Larivi_re_2015], which means that our findings are not representative of the general population of SSH journals. Likewise, this limitation could be extended to our findings in general that whilst not representative of the population of hybrid OA publishing, they provide valuable insight into a key player of scholarly publishing.

Finally, OA policies in favor of hybrid OA in combination with invoicing agreements with Elsevier might be a driver. For instance, we found that nearly half of articles in arts and humanities were facilitated through agreements, which was above-average. This is in line with previous research, suggesting that OA policies and funder mandates have a stronger effect on OA uptake than disciplinary publication cultures [@Huang_2020; @Larivi_re_2018].

Our overall investigation of invoice channels and recipients of hybrid OA APCs found that the majority of Elsevier hybrid articles were invoiced to the author (n=41,725; 58.2%). However, it is important to note that although Elsevier's metadata classifies the sponsor type as "author", this does not necessarily mean that authors paid for APCs out of their own pocket but rather just that the APC was not invoiced through an agreement. Hence, it is possible that APCs invoiced to the author were covered by institutional funds or as part of research grants. For instance, a recent Springer Nature survey found that hybrid OA was predominantly supported through institutional and funder sources (71%), followed by OA agreements (34%), while only 6% were paid from personal funds or savings [@Monaghan_2020]. We also observed notable differences in licensing among invoice types. The majority of hybrid OA articles, regardless of the invoice type, were licensed under the more restrictive CC BY-NC-ND license. There is a lack of dedicated studies on licence selection, but indications so far suggest that when given a choice authors tend to select more restrictive variants for their published works [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013]. On the other hand, when hybrid articles were invoiced through agreements, most articles were licensed under the more liberal CC BY license. As several funding bodies mandate CC BY licenses, including the UK research funders that account for 62% of our "agreement" subsample, this result is perhaps not surprising but suggests the effectiveness of such agreements.

Based on our findings, around a third of the articles were invoiced directly through OA agreements (e.g., research funders, national library consortia). Strikingly, sponsors based in the UK accounted for more than half of these , and a quarter was charged to sponsors from other European countries. This finding reiterates reports from previous studies that the UK's OA profile differs markedly from other countries (cf. ref, ref) and points to the effects of science policy. To increase OA to research publications, the UK implemented centralized APC funding and embraced hybrid OA as a transition model. Within five years of the publication of the Finch Report in 2012, the UK recorded an 18% increase in immediate OA, coupled with a rise in hybrid OA from 2.7% in 2012 to 15.4% of all articles in 2016 [@Jubb_2017], which can be attributed to the availability of APC funding from Research Councils UK (RCUK). Indeed, RCUK block grants have been the largest single source of OA funds in the UK—the Wellcome Trust, and more recently transformative agreements [@Tickel_2018; @Jubb_2017]. However, since then publishing expenditures of UK universities have been rising rapidly (ref), so that several universities stopped supporting hybrid OA in late 2018-2019 [@oxford_2019; @walker_2019; @Castro_2019; @birmingham]. This development might explain the slight decrease in hybrid articles invoiced to the UK our data showed around that time (see Figure \ref{fig:invoice_sponsor_country}).

An essential part in the absence of transparency in scholarly publishing and especially in hybrid OA has been the lack of transparency in publishers’ article management. This has prohibited consumers to get a comprehensive and comparative overview of their publishing activities and the associated costs and, thus, limited their negotiating power. However, certain initiatives in Europe like the ESAC Initiative have begun to press for improved and more transparent workflows for transformative agreements [@Geschuhn_2017]. In particular, the ESAC Workflow Recommendations seek to increase efficiency and transparency in the publication workflow through improved invoicing and reporting processes and metadata delivery, including article-based machine readable information about the OA sponsor. Through this study we demonstrated the utility and benefits of such  publisher-provided metadata about hybrid OA funding and highlighted the need for extending these to include comprehensive information about the licensing and the particular OA type.

Usually identifying delayed OA articles and distinguishing them from other types of OA is challenging since there is no publicly available list of such journals. For instance, Taylor & Francis’ (2021) website indicates that some of their journals participate in a delayed or promotional OA program but no further information is provided as to which journals these are. However, Elsevier provides a list of journals enrolled in the "Open Archive" programme and further tags delayed OA articles in the machine-readable XML version of the article. This enables easy identification and more accurate assessments of hybrid OA, which allows to take into account the substantial amount of delayed OA content in future hybrid OA studies using Elsevier data.

While this study advances our knowledge about hybrid OA uptake and invoicing, the limitations leave room for future research. This study focused on only one publisher, which limits the generalizability of our findings since Elsevier’s attributes such as composition of the journal portfolio, mix of business models, pricing, and promotion of various options cannot be assumed to be representative of scholarly journal publishing in general. Though the invoicing data extracted from the XML files enables unprecedented accuracy and comprehensiveness in the study of hybrid OA, it likely does not disclose the actual OA sponsor in all cases. Most articles were invoiced to authors, but it remains unclear how often authors themselves pay for APCs or invoices are covered through institutional OA funds or research grants. Further research into this topic would increase our understanding of OA funding streams outside of publishing agreements. Further, our study demonstrates that to achieve comprehensive mapping and understanding of the financial flows involved in OA funding, complex, in-depth country-specific meso-level analyses is needed. Such studies would ideally draw on various data sources, such as publicly available bibliometrics and metadata, information about significant national research funders and their OA policies, information about the national library and possible consortia landscape including their publishing agreements, which might require Freedom of Information Requests.

This study provides a snapshot of hybrid OA for the largest journal publisher prior to the impact of the implementation of Plan S requirements. While many Plan S signatories have already had strong OA policies, this harmonized approach is likely to affect publishing decisions of funded authors,the licencing of their articles, and the offerings and pricing of publishers. Since the new requirements are applicable to research funded from 2021 onwards, a comparative study focused on articles  invoiced to Plan S signatories would be a fruitful endeavor.

In a similar vein, future studies could  combine invoice information with research funding information since these contribute different pieces to the funding stream puzzle. This could also provide more insight into the actual APC funding streams behind the author-invoiced articles.

# Conclusion

The primary aim of this empirical study was to investigate Elsevier’s hybrid OA publishing from 2015-2019 to better understand the volume and invoicing of hybrid OA and present a novel, data-driven approach for such analyses. Our results indicate that although the number of hybrid OA articles has increased over time, hybrid OA uptake has remained low. Notably, hybrid APCs were most often invoiced directly to the authors or through institutional agreements, where only a few funding bodies were the main driver of hybrid OA. Finally, our findings highlight that publisher-provided metadata about the invoicing channels  of (hybrid) OA can facilitate research into and increase the understanding of the financial flows of OA publishing.

Ever since its inception, hybrid OA has been a challenging subject to study due to the lack of standardised ways for publishers to flag such content and APC funding data being limited to studies based on surveys and articles’ acknowledgements sections. This study presented a novel approach to studying OA APC invoicing  that is based on publicly available publisher-provided metadata, which can be used on on its own or in combination with other public data sources to gain more detailed and comprehensive insights into hybrid OA uptake and invoicing than ever before. If more publishers reported OA invoicing data on the article level and in a machine-readable format, this would increase transparency and enable  enhanced monitoring of the scholarly journal landscape over time. As hybrid OA has increasingly become a central element of OA policies of research funders and libraries,  customer organisations could require that invoicing information is added to the article-level metadata. As long as publishers do not provide this data in a structured and comprehensive format, they prevent consumers from benchmarking prices and therefore hinder price competition as the only alternative to gain comparative insight is self-reported APC expenditures that research funders and libraries voluntarily report to the Open APC initiative. The Open APC initiative is a valuable service but it cannot replace publisher-provided data since global participation is limited.

From recent science policy developments in Europe it appears that Big Deals have gained support and remain firmly in place in the form of transformative agreements. However, in addition to research funders and libraries that increasingly make publishing contracts public, it is on the publisher to provide  metadata as presented here to improve the transparency in scholarly publishing by linking invoicing data to bibliometrics. From this study we can affirm that hybrid OA is complex as the financial flows come from various stakeholders and invoicing forms, it is not just libraries or consortia, but authors and funders are also directly involved. As of right now, providing  machine-readable article-level invoicing data is not common amongst publishers, which limits the potential for broader studies.

# Data Availability

All data and code used for this study is openly available on GitHub, including version history:

<https://github.com/njahn82/elsevier_hybrid_invoicing>

Formal analyis is written in R Markdown and thus replicable.

# References
